### 2014-05-09 remy_limits.tex
Refine end of Remy chapter

The use of computer-generated congestion-control algorithms presents
different engineering challenges than conventional designs. In
practice, RemyCCs \emph{increase} endpoint complexity in order to
\emph{reduce} the complexity of overall network behavior.

Traditional TCP congestion control specifies simpler behavior for each
endpoint, but the resulting emergent behavior of a multiuser network
is not easily specified and is often suboptimal and variable, and even
unstable. By contrast, our approach focuses on maximizing a
well-specified overall objective at the cost of complex endpoint
algorithms.

We think this tradeoff is ultimately worth taking: today's endpoints
can execute complex algorithms almost as easily as simple ones (and
with Remy, the bulk of the intelligence is computed offline). What
users and system designers ultimately care about, we believe, is the
quality and consistency of overall behavior.

But when endpoint algorithms become very complex, figuring out
\emph{why} a particular RemyCC works involves a reverse-engineering
challenge. To help with this task, I built a visualization and
debugging tool for RemyCCs, called Ratatouille.

Ratatouille displays real-time output from the Remy network simulator,
and plots the current number of packets-in-flight for each concurrent
flow as it evolves over time. The user interacts with a physical USB
control surface---originally intended to control a software music
mixing application---to vary the network parameters: the number of
flows with offered load, the link rate, propagation delay, and buffer
size at the bottleneck queue. The user can then observe how the flows
respond to the changing conditions. (For comparison, Ratatouille can
also run a conventional AIMD TCP-like algorithm.)

By exploring the evolving values of the congestion signals at each
sender as the network changes, we have gained some intuition into how
a RemyCC can achieve good throughput and low delay. When the network
is in steady state, the congestion signals of a RemyCC sender tend to
fall into an orbit, or limit cycle, in which the average number of
packets in flight over the cycle is close to the ``ideal'' number: the
bandwidth-delay product of the bottleneck link divided by the degree
of multiplexing.

The period of the limit cycles is generally a small number of
minRTTs. By contrast, TCP's congestion window typically oscillates on
a much longer timescale that depends on the buffer size, because the
``multiplicative decrease'' feature of TCP's control rules is only
triggered when packet loss is detected (in other words, only after
buffer overflow occurs in the bottleneck queue).

This suggests that a RemyCC's ability to achieve better throughput
than TCP on finite-length flows is mostly due to its quicker response
to changes in conditions---a RemyCC will generally ``ramp up'' to the
ideal number of packets-in-flight more quickly than the TCP
slow-start algorithm.

>>>>

In our experience, high-performing computer-generated
congestion-control algorithms present a different kind of complexity
than existing mechanisms. Traditional TCP congestion control schemes
specify relatively simple rules for each endpoint, but the resulting
emergent behavior of a multiuser network is not easily specified and
can be suboptimal and unstable. By contrast, Remy's algorithms specify
hundreds of rules at the endpoints, but produce more consistent
overall behavior.

We think this tradeoff is ultimately worth taking: today's endpoints
can execute complex algorithms almost as easily as simple ones, and
with Remy, the bulk of the intelligence is computed offline.

But when endpoint algorithms become so complex, it is challenging to
explain \emph{why} a particular RemyCC works without considerable
effort spent reverse-engineering it. To help with this task, I built a
visualization and debugging tool for RemyCCs, called Ratatouille.

Ratatouille displays real-time output from the Remy network simulator,
and plots the current number of packets-in-flight for each concurrent
flow as it evolves over time. The user adjusts a physical USB control
surface (a software mixing board) to vary network parameters,
including the number of flows with offered load and the link rate,
propagation delay, and buffer size at the bottleneck queue. The user
can then observe how the flows respond to the changing
conditions. Ratatouille can also run conventional AIMD TCP-like flows
for comparison.

By exploring the evolving values of the congestion signals, we have
gained some intuition into how a RemyCC can achieve good
performance. When the network is in steady state, the congestion
signals of a RemyCC sender tend to fall into an orbit, or limit cycle,
in which the average number of packets-in-flight over the cycle is
close to the ``ideal'' number: the bandwidth-delay product of the
bottleneck link divided by the degree of multiplexing.

Examining the RemyCCs that are successful in the simulations we
evaluated, we found that the period of the limit cycles is generally a
small number of minRTTs. By contrast, TCP's congestion window
generally oscillates on a much longer timescale that depends on the
buffer size, because the ``multiplicative decrease'' feature of TCP's
control rules is only triggered when packet loss is detected (in other
words, only after buffer overflow occurs in the bottleneck queue).

Our observations suggest that a RemyCC's ability to achieve better
throughput than TCP on finite-length flows is mostly due to its
quicker response to changes in conditions---a RemyCC will generally
``ramp up'' to the ideal number of packets-in-flight more quickly than
the TCP slow-start algorithm.
